# Effect of phenotypic plasticity on subsequent evolutionary dynamics

## Overview

```{r}
total_updates <- 200000
replicates <- 200

all_traits <- c("not","nand","and","ornot","or","andnot")
traits_set_a <- c("not", "and", "or")
traits_set_b <- c("nand", "ornot", "andnot")

# Relative location of data.
working_directory <- "experiments/2021-01-12-evo-dynamics/analysis/" # << For bookdown
# working_directory <- "./"                                              # << For local analysis
```

## Analysis dependencies

Load all required R libraries.

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(Hmisc)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were conducted/knitted with the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
summary_data_loc <- paste0(working_directory, "data/aggregate.csv")
summary_data <- read.csv(summary_data_loc, na.strings="NONE")

summary_data$DISABLE_REACTION_SENSORS <- as.factor(summary_data$DISABLE_REACTION_SENSORS)
summary_data$chg_env <- summary_data$chg_env == "True"
summary_data$dominant_plastic_odd_even <- as.factor(summary_data$dominant_plastic_odd_even)
summary_data$sensors <- summary_data$DISABLE_REACTION_SENSORS == "0"
summary_data$is_plastic <- summary_data$dominant_plastic_odd_even == "True"

env_label_fun <- function(chg_env) {
  if (chg_env) {
    return("Fluctuating")
  } else {
    return("Constant")
  }
}

sensors_label_fun <- function(has_sensors) {
  if (has_sensors) {
    return("Sensors")
  } else {
    return("No sensors")
  }
}

# note that this labeler makes assumptions about how we set up our experiment
condition_label_fun <- function(has_sensors, env_chg) {
  if (has_sensors && env_chg) {
    return("Plastic (fluctuating)")
  } else if (env_chg) {
    return("Non-plastic (fluctuating)")
  } else {
    return("Non-plastic (constant)")
  }
}

summary_data$env_label <- mapply(
  env_label_fun,
  summary_data$chg_env
)
summary_data$sensors_label <- mapply(
  sensors_label_fun,
  summary_data$sensors
)
summary_data$condition <- mapply(
  condition_label_fun,
  summary_data$sensors,
  summary_data$chg_env
)

condition_order = c(
  "Non-plastic (constant)",
  "Non-plastic (fluctuating)",
  "Plastic (fluctuating)"
)

###### time series #####
# time_series_data_loc <- paste0(working_directory, "data/time_series_u0-u200000.csv")
time_series_data_loc <- paste0(working_directory, "data/time_series_u10000-u20000.csv")

time_series_data <- read.csv(time_series_data_loc)
time_series_data$DISABLE_REACTION_SENSORS <- as.factor(time_series_data$DISABLE_REACTION_SENSORS)
time_series_data$chg_env <- time_series_data$chg_env == "True"
time_series_data$sensors <- time_series_data$DISABLE_REACTION_SENSORS == "0"

time_series_data$env_label <- mapply(
  env_label_fun,
  time_series_data$chg_env
)
time_series_data$sensors_label <- mapply(
  sensors_label_fun,
  time_series_data$sensors
)
time_series_data$condition <- mapply(
  condition_label_fun,
  time_series_data$sensors,
  time_series_data$chg_env
)

####### misc #######
# Configure our default graphing theme
theme_set(theme_cowplot())
dir.create(paste0(working_directory, "plots"), showWarnings=FALSE)
```

## Evolution of phenotypic plasticity

For sensor-enabled populations in fluctuating environments, we only transfered populations containing an optimally plastic genotype to phase-two.

```{r}
summary_data_grouped = dplyr::group_by(summary_data, sensors, env_label, condition)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())

ggplot(summary_data_group_counts, aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  geom_text(aes(label=n, y=n+2)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Number of replicates in phase two") +
  theme(
    legend.position="none"
  )
```

We can confirm our expectation that the dominant genotypes in non-plastic conditions are not phenotypically plastic.

```{r}
summary_data_grouped = dplyr::group_by(summary_data, condition, is_plastic)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())
ggplot(filter(summary_data_group_counts, is_plastic), aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  geom_text(aes(label=n, y=n+1)) +
  ylab("Number of replicates with a plastic dominant genotype") +
  theme(
    legend.position="none"
  )
```

## Phenotypic volatility along dominant lineage

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_phenotypic_volatility, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Phenotypic volatility") +
  theme(
    legend.position="none"
  )
```

```{r}
kruskal.test(
  formula=dominant_lineage_phenotypic_volatility~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_phenotypic_volatility,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)

median(filter(summary_data, condition=="Plastic (fluctuating)")$dominant_lineage_phenotypic_volatility)
median(filter(summary_data, condition=="Non-plastic (constant)")$dominant_lineage_phenotypic_volatility)
median(filter(summary_data, condition=="Non-plastic (fluctuating)")$dominant_lineage_phenotypic_volatility)
```

## Mutation accumulation along dominant lineage

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_total_mut_cnt, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Mutation accumulation") +
  theme(
    legend.position="none"
  )
```

```{r}
kruskal.test(
  formula=dominant_lineage_total_mut_cnt~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_total_mut_cnt,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)

median(filter(summary_data, condition=="Plastic (fluctuating)")$dominant_lineage_total_mut_cnt)
median(filter(summary_data, condition=="Non-plastic (constant)")$dominant_lineage_total_mut_cnt)
median(filter(summary_data, condition=="Non-plastic (fluctuating)")$dominant_lineage_total_mut_cnt)
```

### Mutation accumulation normalized by lineage length

```{r}
summary_data$mutations_per_lineage_step <- summary_data$dominant_lineage_total_mut_cnt / summary_data$dominant_lineage_length_genotypes
ggplot(summary_data, aes(x=condition, y=mutations_per_lineage_step, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Mutation accumulation / lineage length") +
  theme(
    legend.position="none"
  )
```

```{r}
kruskal.test(
  formula=mutations_per_lineage_step~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$mutations_per_lineage_step,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)

median(filter(summary_data, condition=="Plastic (fluctuating)")$mutations_per_lineage_step)
median(filter(summary_data, condition=="Non-plastic (constant)")$mutations_per_lineage_step)
median(filter(summary_data, condition=="Non-plastic (fluctuating)")$mutations_per_lineage_step)
```

## Selective sweeps

The number of times the most recent common ancestor changes gives us the number of selective sweeps that occur during the experiment.

```{r}
ggplot(summary_data, aes(x=condition, y=phylo_mrca_changes, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Number of selective sweeps") +
  theme(
    legend.position="none"
  )
```

```{r}
kruskal.test(
  formula=phylo_mrca_changes~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$phylo_mrca_changes,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)

median(filter(summary_data, condition=="Plastic (fluctuating)")$phylo_mrca_changes)
median(filter(summary_data, condition=="Non-plastic (constant)")$phylo_mrca_changes)
median(filter(summary_data, condition=="Non-plastic (fluctuating)")$phylo_mrca_changes)
```

## Genome length

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_genome_length, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Genome length") +
  theme(
    legend.position="none"
  )
```

## Diversity over time

### Shannon entropy over time

```{r}
# shannon entropy
ggplot(time_series_data, aes(x=update, y=phylo_diversity, fill=condition, color=condition)) +
  stat_summary(fun="mean", geom="line") +
  stat_summary(
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    geom="ribbon",
    alpha=0.2,
    linetype=0
  ) +
  ylab("Shannon entropy")
```

Zoom in to just 400 update span.

```{r}
# shannon entropy
ggplot(filter(time_series_data, update <= 10400), aes(x=update, y=phylo_diversity, fill=condition, color=condition)) +
  stat_summary(fun="mean", geom="line") +
  stat_summary(
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    geom="ribbon",
    alpha=0.2,
    linetype=0
  ) +
  ylab("Shannon entropy")
```

### Phylogenetic diversity over time

```
// From (Faith 1992, reviewed in Winters et al., 2013), phylogenetic diversity is the sum of edges in the minimal spanning tree connected the taxa you're calculating diversity of.
// This calculates phylogenetic diversity for all extant taxa in the tree.
```

```{r}
ggplot(time_series_data, aes(x=update, y=phylo_current_phylogenetic_diversity, fill=condition, color=condition)) +
  stat_summary(fun="mean", geom="line") +
  stat_summary(
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    geom="ribbon",
    alpha=0.2,
    linetype=0
  ) +
  ylab("Phylogenetic diversity")
```

### Number of extant taxa (genotypes) over time

```{r}
ggplot(time_series_data, aes(x=update, y=phylo_num_taxa_extant, fill=condition, color=condition)) +
  stat_summary(fun="mean", geom="line") +
  stat_summary(
    fun.data="mean_cl_boot",
    fun.args=list(conf.int=0.95),
    geom="ribbon",
    alpha=0.2,
    linetype=0
  )
```

## Camera-ready figures

Figures styled for the paper.


Mutation accumulation

```{r}
library(viridis)
library(RColorBrewer)
# dominant_lineage_total_mut_cnt or mutations_per_lineage_step?
mutation_count_fig <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_total_mut_cnt, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=c(
      "Constant env.",
      "Fluct. env.\nnon-plastic",
      "Fluct. env.\nplastic"
    )
  ) +
  scale_y_continuous(
    name="Mutations accumulated (log scale)",
    trans="log10"
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip() +
  ggsave(
    paste0(working_directory, "plots/", "mutation-accumulation.pdf"),
    width=5,
    height=4
  )
```

Phenotypic volatility

```{r}
phenotypic_volatility_fig <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_phenotypic_volatility, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=c(
      "Constant env.",
      "Fluct. env.\nnon-plastic",
      "Fluct. env.\nplastic"
    )
  ) +
  scale_y_continuous(
    name="Phenotypic volatility (log scale)",
    trans="log10",
    limits=c(50,11000)
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none",
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.title.y=element_blank()
  ) +
  coord_flip() +
  ggsave(
    paste0(working_directory, "plots/", "phenotypic-volatility.pdf"),
    width=4,
    height=4
  )
```

Selective sweeps.

```{r}
selective_sweeps_fig <- ggplot(
    summary_data,
    aes(x=condition, y=phylo_mrca_changes, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=c(
      "Constant env.",
      "Fluct. env.\nnon-plastic",
      "Fluct. env.\nplastic"
    )
  ) +
  scale_y_continuous(
    name="Selective sweeps (log scale)",
    trans="log10"
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none",
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.title.y=element_blank()
  ) +
  coord_flip() +
  ggsave(
    paste0(working_directory, "plots/", "selective-sweeps.pdf"),
    width=4,
    height=4
  )

# nrow(filter(summary_data, condition=="Non-plastic (fluctuating)" & phylo_mrca_changes <= 1))
```

All together

```{r}
grid <- plot_grid(
  mutation_count_fig,
  phenotypic_volatility_fig,
  selective_sweeps_fig,
  nrow=1,
  align="v",
  labels="auto"
)
save_plot(
   paste0(working_directory, "plots/", "evolutionary-dynamics.pdf"),
   grid,
   base_height=6,
   base_asp=2.5
)
# ggsave(
#   paste0(working_directory, "plots/", "evolutionary-dynamics.pdf"),
#   width=10,
#   height=8
# )

```