# Evolutionary change

The effect of adaptive phenotypic plasticity on evolutionary change.

## Overview

```{r}
total_updates <- 200000
replicates <- 100

all_traits <- c("not","nand","and","ornot","or","andnot")
traits_set_a <- c("not", "and", "or")
traits_set_b <- c("nand", "ornot", "andnot")

# Relative location of data.
working_directory <- "experiments/2021-02-08-evo-dynamics/analysis/" # << For bookdown
# working_directory <- "./"                                              # << For local analysis
```

## Analysis dependencies

Load all required R libraries.

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(Hmisc)
library(boot)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were conducted/knitted with the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
summary_data_loc <- paste0(working_directory, "data/aggregate.csv")
summary_data <- read.csv(summary_data_loc, na.strings="NONE")

summary_data$DISABLE_REACTION_SENSORS <- as.factor(summary_data$DISABLE_REACTION_SENSORS)
summary_data$chg_env <- summary_data$chg_env == "True"
summary_data$dominant_plastic_odd_even <- as.factor(summary_data$dominant_plastic_odd_even)
summary_data$sensors <- summary_data$DISABLE_REACTION_SENSORS == "0"
summary_data$is_plastic <- summary_data$dominant_plastic_odd_even == "True"

env_label_fun <- function(chg_env) {
  if (chg_env) {
    return("Fluctuating")
  } else {
    return("Constant")
  }
}

sensors_label_fun <- function(has_sensors) {
  if (has_sensors) {
    return("Sensors")
  } else {
    return("No sensors")
  }
}

# note that this labeler makes assumptions about how we set up our experiment
condition_label_fun <- function(has_sensors, env_chg) {
  if (has_sensors && env_chg) {
    return("PLASTIC")
  } else if (env_chg) {
    return("NON-PLASTIC")
  } else {
    return("STATIC")
  }
}

summary_data$env_label <- mapply(
  env_label_fun,
  summary_data$chg_env
)
summary_data$sensors_label <- mapply(
  sensors_label_fun,
  summary_data$sensors
)
summary_data$condition <- mapply(
  condition_label_fun,
  summary_data$sensors,
  summary_data$chg_env
)

condition_order = c(
  "STATIC",
  "NON-PLASTIC",
  "PLASTIC"
)

####### misc #######
# Configure our default graphing theme
theme_set(theme_cowplot())
dir.create(paste0(working_directory, "plots"), showWarnings=FALSE)
```

## Evolution of phenotypic plasticity

For sensor-enabled populations in fluctuating environments, we only transfered populations containing an optimally plastic genotype to phase-two.

```{r}
summary_data_grouped = dplyr::group_by(summary_data, condition)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())

ggplot(summary_data_group_counts, aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  geom_text(aes(label=n, y=n+2)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Number of replicates in phase two") +
  theme(
    legend.position="none"
  )
```

We can confirm our expectation that the dominant genotypes in non-plastic conditions are not phenotypically plastic.

```{r, message=FALSE}
summary_data_grouped = dplyr::group_by(summary_data, condition, is_plastic)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())
ggplot(filter(summary_data_group_counts, is_plastic), aes(x=condition, y=n, fill=condition)) +
  geom_col(
    position=position_dodge(0.9)
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  geom_text(aes(label=n, y=n+1)) +
  ylab("Number of plastic replicates") +
  ylim(0, 100) +
  theme(
    legend.position="none"
  )
```

## Average generation

```{r}
ggplot(summary_data, aes(x=condition, y=time_average_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("average generation") +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$time_average_generation)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$time_average_generation)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$time_average_generation)
)

kruskal.test(
  formula=time_average_generation~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$time_average_generation,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

## Selective sweeps

The number of times the most recent common ancestor changes gives us the number of selective sweeps that occur during the experiment.

```{r}
ggplot(summary_data, aes(x=condition, y=phylo_mrca_changes, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Number of selective sweeps") +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$phylo_mrca_changes)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$phylo_mrca_changes)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$phylo_mrca_changes)
)

kruskal.test(
  formula=phylo_mrca_changes~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$phylo_mrca_changes,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Average number of generations between selective sweeps

```{r}
summary_data$generations_per_mrca_change <- summary_data$time_average_generation / summary_data$phylo_mrca_changes

ggplot(summary_data, aes(x=condition, y=generations_per_mrca_change, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$generations_per_mrca_change)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$generations_per_mrca_change)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$generations_per_mrca_change)
)

kruskal.test(
  formula=generations_per_mrca_change~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$generations_per_mrca_change,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

## Phenotypic volatility along dominant lineage

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_trait_volatility, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_y_continuous(
    name="Phenotypic volatility (log scale)",
    trans="pseudo_log",
    breaks=c(0, 10, 100, 1000, 10000),
    limits=c(-1,10000)
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_trait_volatility)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_trait_volatility)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_trait_volatility)
)

kruskal.test(
  formula=dominant_lineage_trait_volatility~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_trait_volatility,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Phenotypic volatility normalized by generations elapsed

```{r}
summary_data$dominant_lineage_trait_volatility_per_generation <- summary_data$dominant_lineage_trait_volatility / summary_data$dominant_generation_born

ggplot(summary_data, aes(x=condition, y=dominant_lineage_trait_volatility_per_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_trait_volatility_per_generation)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_trait_volatility_per_generation)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_trait_volatility_per_generation)
)

kruskal.test(
  formula=dominant_lineage_trait_volatility_per_generation~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_trait_volatility_per_generation,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Phenotypic volatility normalized by lineage length

Lineage length = number of genotypes along the lineage.

```{r}
summary_data$dominant_lineage_trait_volatility_per_lineage_step <- summary_data$dominant_lineage_trait_volatility / summary_data$dominant_lineage_length_genotypes

ggplot(summary_data, aes(x=condition, y=dominant_lineage_trait_volatility_per_lineage_step, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_trait_volatility_per_lineage_step)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_trait_volatility_per_lineage_step)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_trait_volatility_per_lineage_step)
)

kruskal.test(
  formula=dominant_lineage_trait_volatility_per_lineage_step~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_trait_volatility_per_lineage_step,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Phenotypic fidelity

Frequency that an offspring's genotype is identical to a parent genotype (along the dominant lineage).

```{r}
summary_data$dominant_lineage_trait_fidelity <- (summary_data$dominant_generation_born - summary_data$dominant_lineage_trait_volatility) / summary_data$dominant_generation_born

ggplot(summary_data, aes(x=condition, y=dominant_lineage_trait_fidelity, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_trait_fidelity)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_trait_fidelity)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_trait_fidelity)
)

kruskal.test(
  formula=dominant_lineage_trait_fidelity~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_trait_fidelity,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

## Mutation accumulation along the dominant lineage

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_total_mut_cnt, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Mutation accumulation") +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_total_mut_cnt)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_total_mut_cnt)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_total_mut_cnt)
)

kruskal.test(
  formula=dominant_lineage_total_mut_cnt~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_total_mut_cnt,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Mutation accumulation normalized by generations elapsed

```{r}
summary_data$mutations_per_generation <- summary_data$dominant_lineage_total_mut_cnt / summary_data$dominant_generation_born

ggplot(summary_data, aes(x=condition, y=mutations_per_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Mutation accumulation / generation") +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$mutations_per_generation)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$mutations_per_generation)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$mutations_per_generation)
)

kruskal.test(
  formula=mutations_per_generation~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$mutations_per_generation,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Mutation accumulation normalized by lineage length

```{r}
summary_data$mutations_per_lineage_step <- summary_data$dominant_lineage_total_mut_cnt / summary_data$dominant_lineage_length_genotypes
ggplot(summary_data, aes(x=condition, y=mutations_per_lineage_step, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Mutation accumulation / lineage length") +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$mutations_per_lineage_step)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$mutations_per_lineage_step)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$mutations_per_lineage_step)
)

kruskal.test(
  formula=mutations_per_lineage_step~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$mutations_per_lineage_step,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Genotypic fidelity

The frequency that an offspring's genotype is the same as a parent's genotype.

```{r}
summary_data$dominant_lineage_genotypic_fidelity <- (summary_data$dominant_generation_born - summary_data$dominant_lineage_num_mut_steps) / summary_data$dominant_generation_born

ggplot(summary_data, aes(x=condition, y=dominant_lineage_genotypic_fidelity, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_genotypic_fidelity)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_genotypic_fidelity)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_genotypic_fidelity)
)

kruskal.test(
  formula=dominant_lineage_genotypic_fidelity~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$dominant_lineage_genotypic_fidelity,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

### Characterizing variation along lineages

#### How many mutation-steps along the lineage result in phenotypic changes?

```{r}
summary_data$frac_phenotype_changing_mut_steps <- summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype / summary_data$dominant_lineage_num_mut_steps

ggplot(filter(summary_data, dominant_lineage_num_mut_steps > 0), aes(x=frac_phenotype_changing_mut_steps)) +
  geom_density() +
  facet_wrap(
    ~condition,
    nrow=3,
    labeller=label_both

  ) +
  xlim(0, 1.0) +
  theme(
    legend.position="none"
  )
```

```{r}
ggplot(summary_data, aes(x=condition, y=frac_phenotype_changing_mut_steps, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC: ",
  median(filter(summary_data, condition=="PLASTIC")$frac_phenotype_changing_mut_steps)
)
paste0(
  "STATIC: ",
  median(filter(summary_data, condition=="STATIC")$frac_phenotype_changing_mut_steps)
)
paste0(
  "NON-PLASTIC: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$frac_phenotype_changing_mut_steps)
)

kruskal.test(
  formula=frac_phenotype_changing_mut_steps~condition,
  data=summary_data
)

pairwise.wilcox.test(
  x=summary_data$frac_phenotype_changing_mut_steps,
  g=summary_data$condition,
  p.adjust.method="bonferroni",
)
```

```{r, include=FALSE}
samplemean <- function(x, d) {
  return(mean(x[d]))
}

# for (cond in condition_order) {
#   # How many mutational steps result in a change in aggregate phenotype?
#   print(paste0(cond, " - Mean with bootstrapped 95% CI"))
#   bo <- boot(filter(summary_data, condition==cond & dominant_lineage_num_mut_steps > 0)$frac_phenotype_changing_mut_steps, statistic=samplemean, R=10000)
#   print(bo)
#   print(boot.ci(bo, conf=0.95, type="perc"))
# }
```

#### For PLASTIC populations, what fraction of phenotype-altering mutations affect the unexpressed phenotype?

```{r}
summary_data$frac_unexpressed_mut_steps <- summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype / summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype
summary_data$frac_expressed_mut_steps <- summary_data$dominant_lineage_num_mut_steps_that_change_expressed_phenotype / summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype

ggplot(filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_aggregate_phenotype > 0), aes(x=frac_unexpressed_mut_steps)) +
  geom_density() +
  theme(
    legend.position="none"
  )

print(paste0("PLASTIC - Mean with bootstrapped 95% CI"))
bo <- boot(filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_aggregate_phenotype > 0)$frac_unexpressed_mut_steps, statistic=samplemean, R=10000)
print(bo)
print(boot.ci(bo, conf=0.95, type="perc"))
```

```{r}
plastic_summary_data <- filter(summary_data, condition=="PLASTIC")
aggregate_frac_mut_steps_that_change_unexpressed_phenotype <- sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype) / sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype)
sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype)
sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype)
aggregate_frac_mut_steps_that_change_unexpressed_phenotype
```

`r sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype)` / `r sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_aggregate_phenotype)` (`r aggregate_frac_mut_steps_that_change_unexpressed_phenotype`)

#### For PLASTIC populations, what fraction of mutations that affect the unexpressed phenotype are deleterious versus beneficial?


```{r}
aggregate_frac_unexpressed_deleterious_mut_steps <- sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype_deleterious) / sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype)
aggregate_frac_unexpressed_beneficial_mut_steps <- sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype_beneficial) / sum(plastic_summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype)
```


Deleterious

```{r}

summary_data$frac_unexpressed_deleterious_mut_steps <- summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype_deleterious / summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype

ggplot(
  filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype > 0),
  aes(x=frac_unexpressed_deleterious_mut_steps)
  ) +
  geom_density() +
  theme(
    legend.position="none"
  )

bo <- boot(filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_aggregate_phenotype > 0)$frac_unexpressed_deleterious_mut_steps, statistic=samplemean, R=10000)
print(bo)
print(boot.ci(bo, conf=0.95, type="perc"))
```

Beneficial

```{r}
summary_data$frac_unexpressed_beneficial_mut_steps <- summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype_beneficial / summary_data$dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype

ggplot(
  filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_unexpressed_phenotype > 0),
  aes(x=frac_unexpressed_beneficial_mut_steps)
  ) +
  geom_density() +
  theme(
    legend.position="none"
  )

bo <- boot(filter(summary_data, condition=="PLASTIC" & dominant_lineage_num_mut_steps_that_change_aggregate_phenotype > 0)$frac_unexpressed_beneficial_mut_steps, statistic=samplemean, R=10000)
print(bo)
print(boot.ci(bo, conf=0.95, type="perc"))
```

## Depth of MRCA

```{r}
ggplot(summary_data, aes(x=condition, y=phylo_mrca_depth, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("MRCA Depth") +
  theme(
    legend.position="none"
  )
```

## Manuscript figures

Figures styled for the paper.

### Evolutionary change panel

Selective sweeps, mutation accumulation, phenotypic volatility.

Mutation accumulation:

```{r}
# dominant_lineage_total_mut_cnt or mutations_per_lineage_step?
mutation_count_fig <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_total_mut_cnt, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Mutations accumulated (log scale)",
    trans="log10",
    breaks=c(100, 1000, 10000),
    limits=c(100, 10000)
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/", "mutation-accumulation.pdf"),
    width=5,
    height=4
  )
mutation_count_fig
```

Phenotypic volatility:

```{r}
phenotypic_volatility_fig <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_trait_volatility, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Phenotypic volatility (log scale)",
    trans="pseudo_log",
    breaks=c(0, 10, 100, 1000, 10000),
    limits=c(-1,10000)
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/", "phenotypic-volatility.pdf"),
    width=4,
    height=4
  )

phenotypic_volatility_fig
```

Selective sweeps:

```{r}
selective_sweeps_fig <- ggplot(
    summary_data,
    aes(x=condition, y=phylo_mrca_changes, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Coalescence Events (log scale)",
    trans="log10",
    breaks=c(10, 100, 1000),
    limits=c(10, 1000)
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/", "selective-sweeps.pdf"),
    width=4,
    height=4
  )

selective_sweeps_fig
```

All together:

```{r}
grid <- plot_grid(
  selective_sweeps_fig + theme(
    legend.position="none"
  ),
  mutation_count_fig + theme(
    legend.position="none",
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.title.y=element_blank()
  ),
  phenotypic_volatility_fig + theme(
    legend.position="none",
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank(),
    axis.title.y=element_blank()
  ),
  nrow=1,
  align="v",
  labels="auto"
)
grid
save_plot(
   paste0(working_directory, "plots/", "evolutionary-dynamics.pdf"),
   grid,
   base_height=6,
   base_asp=2.5
)
```