---
title: "2020-08-28 Experiment Analysis"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)   # (Wickham, 2016)
library(tidyr)     # (Wickham and Henry, 2020)
library(dplyr)     # (Wickham et al., 2020)
library(reshape2)  # (Wickham, 2007)
library(cowplot)   # (Wilke, 2019)
library(patchwork) # (Pederson, 2020)
library(viridis)   # (Garnier, 2018)
library(hexbin)
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
data_path <- "./data/aggregate.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$DISABLE_REACTION_SENSORS <- as.factor(agg_data$DISABLE_REACTION_SENSORS)
agg_data$environment <- factor(
  agg_data$environment,
  levels=c("a-u0", "b-u0", "all-u0",
           "chg-u3", "chg-u10","chg-u30",
           "chg-u100","chg-u300","chg-u1000")
)
agg_data$chg_env <- as.factor(agg_data$chg_env)
agg_data$plastic_odd_even <- as.factor(agg_data$plastic_odd_even)
agg_data$sensors <- agg_data$DISABLE_REACTION_SENSORS == "0"

theme_set(theme_cowplot())
```

## Graphs!

### Plasticity

```{r}
ggplot(filter(agg_data, plastic_odd_even=="True"), aes(x=sensors)) +
  geom_bar(position=position_dodge(0.9)) +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9), vjust=0
  )  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  ylab("plastic final dominant (odd/even environment)") +
  ylim(0, 50) +
  facet_wrap(~environment, nrow=1) +
  ggtitle("Plastic?") +
  ggsave("dom_plastic.pdf", width=10, height=7)
```

```{r}
ggplot(filter(agg_data, optimal_plastic=="True"), aes(x=sensors)) +
  geom_bar(position=position_dodge(0.9)) +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9), vjust=0
  )  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  ylab("plastic final dominant (odd/even environment)") +
  ylim(0, 50) +
  facet_wrap(~environment, nrow=1) +
  ggtitle("Optimal plastic?") +
  ggsave("dom_plastic_optimal.pdf")
```

### Match scores

```{r}
ggplot(filter(agg_data, chg_env=="True"), aes(x=sensors, y=match_score_odd_even)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, nrow=1) 
```

### MRCA Changes

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_mrca_changes, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, nrow=1) +
  ggsave("mrca_changes.pdf", width=10, height=7)
```

### Lineage Length (genotypes)

```{r}
ggplot(agg_data, aes(x=sensors, y=lineage_length_genotypes, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, nrow=1) +
  # theme(
  #   panel.border = element_rect(colour = "black", fill=NA, size=1)
  # ) +
  ggsave("lineage_length_genotypes.pdf", width=10, height=7)
```

### Mutation Accumulation

```{r}
ggplot(agg_data, aes(x=sensors, y=total_mut_cnt, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, ncol=1, scales="free") +
  ggsave("total_mutations_accumulated.pdf", width=5, height=20)
```

### Phylogenetic Diversity

Current phylogentic diversity = size of phylogentic tree (number of active taxa + ancestral taxa represented in phylogeny).

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_current_phylogenetic_diversity)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, nrow=1) 
```

### Taxa Entropy (diversity)

Extant diversity measured as entropy of taxa in population.

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_diversity)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  facet_wrap(~environment, nrow=1) 
```