# Genetic hitchhiking

The effect of adaptive phenotypic plasticity on (deleterious) genetic hitchhiking.

## Overview

```{r}
total_updates <- 200000
replicates <- 100

focal_traits <- c("not","nand","and","ornot","or","andnot")
traits_set_a <- c("not", "and", "or")
traits_set_b <- c("nand", "ornot", "andnot")

# Relative location of data.
working_directory <- "experiments/2021-02-05-hitchhiking/analysis/" # << For bookdown
# working_directory <- "./"
```

## Analysis dependencies

Load all required R libraries.

```{r, message=FALSE}
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
library(cowplot)
library(Hmisc)
library(boot)
library(fmsb)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were conducted/knitted with the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
####### summary data #######
summary_data_loc <- paste0(working_directory, "data/aggregate.csv")
summary_data <- read.csv(summary_data_loc, na.strings="NONE")

summary_data$DISABLE_REACTION_SENSORS <- as.factor(summary_data$DISABLE_REACTION_SENSORS)
summary_data$chg_env <- summary_data$chg_env == "True"
summary_data$dominant_plastic_odd_even <- as.factor(summary_data$dominant_plastic_odd_even)
summary_data$sensors <- summary_data$DISABLE_REACTION_SENSORS == "0"
summary_data$is_plastic <- summary_data$dominant_plastic_odd_even == "True"
summary_data$POISON_PENALTY <- as.factor(summary_data$POISON_PENALTY)

env_label_fun <- function(chg_env) {
  if (chg_env) {
    return("Fluctuating")
  } else {
    return("Constant")
  }
}

sensors_label_fun <- function(has_sensors) {
  if (has_sensors) {
    return("Sensors")
  } else {
    return("No sensors")
  }
}

condition_label_fun <- function(has_sensors, env_chg) {
  if (has_sensors && env_chg) {
    return("PLASTIC")
  } else if (env_chg) {
    return("NON-PLASTIC")
  } else {
    return("STATIC")
  }
}

summary_data$env_label <- mapply(
  env_label_fun,
  summary_data$chg_env
)
summary_data$sensors_label <- mapply(
  sensors_label_fun,
  summary_data$sensors
)
summary_data$condition <- mapply(
  condition_label_fun,
  summary_data$sensors,
  summary_data$chg_env
)

condition_order = c(
  "STATIC",
  "NON-PLASTIC",
  "PLASTIC"
)

###### time series #####
lineage_time_series_data_loc <- paste0(working_directory, "data/lineage_series.csv")
lineage_time_series_data <- read.csv(lineage_time_series_data_loc)

lineage_time_series_data$DISABLE_REACTION_SENSORS <- as.factor(lineage_time_series_data$DISABLE_REACTION_SENSORS)
lineage_time_series_data$chg_env <- lineage_time_series_data$chg_env == "True"
lineage_time_series_data$sensors <- lineage_time_series_data$DISABLE_REACTION_SENSORS == "0"
lineage_time_series_data$POISON_PENALTY <- as.factor(lineage_time_series_data$POISON_VALUE)

lineage_time_series_data$env_label <- mapply(
  env_label_fun,
  lineage_time_series_data$chg_env
)
lineage_time_series_data$sensors_label <- mapply(
  sensors_label_fun,
  lineage_time_series_data$sensors
)
lineage_time_series_data$condition <- mapply(
  condition_label_fun,
  lineage_time_series_data$sensors,
  lineage_time_series_data$chg_env
)

####### misc #######
# Configure our default graphing theme
theme_set(theme_cowplot())
dir.create(paste0(working_directory, "plots"), showWarnings=FALSE)
```

## Evolution of phenotypic plasticity

For sensor-enabled populations in fluctuating environments, we only transfered populations containing an optimally plastic genotype to phase-two.

```{r, message=FALSE}
summary_data_grouped = dplyr::group_by(summary_data, sensors, env_label, condition, POISON_PENALTY)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())

ggplot(summary_data_group_counts, aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  geom_text(aes(label=n, y=n+2)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Number of replicates in phase two") +
  facet_wrap(~POISON_PENALTY, labeller=label_both) +
  theme(
    legend.position="none"
  )
```

We can confirm our expectation that the dominant genotypes in non-plastic conditions are not phenotypically plastic.

```{r}
summary_data_grouped = dplyr::group_by(summary_data, condition, is_plastic, POISON_PENALTY)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())
ggplot(filter(summary_data_group_counts, is_plastic), aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  geom_text(aes(label=n, y=n+1)) +
  ylab("Number of replicates with a plastic dominant genotype") +
  ylim(0, 100) +
  facet_wrap(~POISON_PENALTY, labeller=label_both) +
  theme(
    legend.position="none"
  )
```

## Hitchhiking instruction execution

### Number of replicates where final dominant genotype executes hitchhiker instruction

```{r}
hitchiker_penalty <- 0.1
```

```{r}
occurrences <- c(
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC" & dominant_times_poison_executed > 0)$RANDOM_SEED),
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="PLASTIC" & dominant_times_poison_executed > 0)$RANDOM_SEED),
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="STATIC" & dominant_times_poison_executed > 0)$RANDOM_SEED)
)
trials <- c(
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC")$RANDOM_SEED),
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="PLASTIC")$RANDOM_SEED),
  length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="STATIC" )$RANDOM_SEED)
)
names(trials) <- c(
  "NON-PLASTIC",
  "PLASTIC",
  "STATIC"
)
names(occurrences) <- c(
  "NON-PLASTIC",
  "PLASTIC",
  "STATIC"
)

pairwise.fisher.test(x=occurrences, n=trials, p.adjust.method="bonferroni")
```

### Final dominant genotype hitchhiker execution

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_times_poison_executed, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("poison instruction executions (final dominant)") +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both,
    scale="free_y"
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-poison.pdf"),
    width=15,
    height=10
  )
```

```{r}
penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=dominant_times_poison_executed~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$dominant_times_poison_executed,
      g=stat_data$condition,
      p.adjust.method="bonferroni"
    )
  )
}
```

### Hitchhiker instruction execution in final population

```{r}
ggplot(summary_data, aes(x=condition, y=final_population_poison, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_y_continuous(
    name="poison instruction executions (final population)",
    trans="pseudo_log",
    breaks=c(0,10,100,1000, 10000, 100000, 1000000),
    limits=c(-1,1000000)
  ) +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/final-population-poison-log.pdf"),
    width=15,
    height=10
  )
```

```{r}
penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=final_population_poison~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$final_population_poison,
      g=stat_data$condition,
      p.adjust.method="bonferroni"
    )
  )
}
```

### Hitchhiker instruction execution along final dominant lineage (cummulative)

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_times_poison_executed, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_y_continuous(
    name="poison instruction executions (dominant lineage)",
    trans="pseudo_log",
    breaks=c(0,10,100,1000,10000,100000),
    limits=c(-1,100000)
  ) +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/final-dominant-lineage-poison-log.pdf"),
    width=15,
    height=10
  )
```

```{r}
penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=dominant_lineage_times_poison_executed~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$dominant_lineage_times_poison_executed,
      g=stat_data$condition,
      p.adjust.method="bonferroni"
    )
  )
}
```

## Characterizing mutations that increase hitchhiker instruction execution

### Number of offspring along dominant lineage with increase in hitchiker instruction execution

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_num_times_hitchhike_inst_exec_increases, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/final-dominant-lineage-poison-increase-num-mutants-log.pdf"),
    width=15,
    height=10
  )
```

```{r}
penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=dominant_lineage_num_times_hitchhike_inst_exec_increases~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$dominant_lineage_num_times_hitchhike_inst_exec_increases,
      g=stat_data$condition,
      p.adjust.method="bonferroni"
    )
  )
}

sum(filter(summary_data, condition=="NON-PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)
sum(filter(summary_data, condition=="PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)
sum(filter(summary_data, condition=="STATIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)

# sum(filter(summary_data, condition=="NON-PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_trait_volatility)
# sum(filter(summary_data, condition=="PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_trait_volatility)
# sum(filter(summary_data, condition=="STATIC" & POISON_PENALTY==0.1)$dominant_lineage_trait_volatility)

```

#### Normalized by generations

```{r}
summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases_per_generation <- summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases / summary_data$dominant_generation_born
ggplot(summary_data, aes(x=condition, y=dominant_lineage_num_times_hitchhike_inst_exec_increases_per_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=dominant_lineage_num_times_hitchhike_inst_exec_increases_per_generation~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$dominant_lineage_num_times_hitchhike_inst_exec_increases_per_generation,
      g=stat_data$condition,
      p.adjust.method="bonferroni"
    )
  )
}
```

### What fraction of mutations that increase hitchhiker instruction execution co-occur with base trait changes?

```{r}
# Fraction of unexpressed vs expressed increases in hitchhiker instructions
summary_data$frac_hitchhiking_linked_trait_change <- summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases_with_primary_trait_change / summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases

ggplot(filter(summary_data, dominant_lineage_num_times_hitchhike_inst_exec_increases>0 ), aes(x=frac_hitchhiking_linked_trait_change)) +
  geom_density() +
  facet_grid(
    condition~POISON_PENALTY,
    labeller=label_both,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-lineage-frac_hitchhiking_linked_trait_change.pdf"),
    width=15,
    height=10
  )

ggplot(filter(summary_data, dominant_lineage_num_times_hitchhike_inst_exec_increases>0 ), aes(x=condition, y=frac_hitchhiking_linked_trait_change, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  facet_wrap(
    ~POISON_PENALTY,
    labeller=label_both,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  )
```

```{r}

penalties <- levels(summary_data$POISON_PENALTY)
for (penalty in penalties) {
  stat_data <- filter(summary_data, POISON_PENALTY==penalty & dominant_lineage_num_times_hitchhike_inst_exec_increases>0)
  print(
    paste0(
      "PENALTY: ", penalty
    )
  )
  kt <- kruskal.test(
      formula=frac_hitchhiking_linked_trait_change~condition,
      data=stat_data
    )
  print(
    kt
  )
  if (is.na(kt$p.value)) { next }
  if (kt$p.value > 0.05) { next }
  print(
    pairwise.wilcox.test(
      x=stat_data$frac_hitchhiking_linked_trait_change,
      g=stat_data$condition,
      p.adjust.method="bonferroni",
      exact=FALSE
    )
  )
}

denom <- sum(filter(summary_data, condition=="NON-PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)
num <- sum(filter(summary_data, condition=="NON-PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases_with_primary_trait_change)
paste0("NON-PLASTIC: ", num/denom, "(", num, "/", denom, ")")

denom <- sum(filter(summary_data, condition=="PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)
num <- sum(filter(summary_data, condition=="PLASTIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases_with_primary_trait_change)
paste0("PLASTIC: ", num/denom, " (", num, "/", denom, ")")

denom <- sum(filter(summary_data, condition=="STATIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases)
num <- sum(filter(summary_data, condition=="STATIC" & POISON_PENALTY==0.1)$dominant_lineage_num_times_hitchhike_inst_exec_increases_with_primary_trait_change)
paste0("STATIC: ", num/denom, " (", num, "/", denom, ")")

```

### What about unexpressed vs expressed trait changes in plastic populations?

```{r}
summary_data$frac_unexpressed_hitchhiker_inc <- summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases_in_unexpressed_phenotype / summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases
summary_data$frac_expressed_hitchiker_inc <- summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases_in_expressed_phenotype / summary_data$dominant_lineage_num_times_hitchhike_inst_exec_increases

ggplot(filter(summary_data, dominant_lineage_num_times_hitchhike_inst_exec_increases>0 & condition=="PLASTIC"), aes(x=frac_unexpressed_hitchhiker_inc)) +
  geom_density() +
  facet_grid(
    condition~POISON_PENALTY,
    labeller=label_both,
    scales="free_y"
  ) +
  theme(
    legend.position="none"
  )

# ggplot(filter(summary_data, dominant_lineage_num_times_hitchhike_inst_exec_increases>0 & condition=="PLASTIC"), aes(x=frac_expressed_hitchiker_inc)) +
#   geom_density() +
#   facet_grid(
#     condition~POISON_PENALTY,
#     labeller=label_both,
#     scales="free_y"
#   ) +
#   theme(
#     legend.position="none"
#   )
```

## Manuscript figures

```{r}
hitchiker_penalty <- 0.1
```

```{r}
ggplot(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC" & dominant_times_poison_executed > 0), aes(x=condition, fill=condition) ) +
  geom_bar() +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9),
    vjust=0
  ) +
  scale_y_continuous(
    name="Number of replicates where poison is executed by dominant genotype",
    limits=c(0, replicates+2)
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order,
    breaks=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  )

fig_data <- data.frame(
  num_execute_poison=c(
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC" & dominant_times_poison_executed > 0)$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="PLASTIC" & dominant_times_poison_executed > 0)$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="STATIC" & dominant_times_poison_executed > 0)$RANDOM_SEED)
  ),
  num_execute_no_poison=c(
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC" & dominant_times_poison_executed == 0)$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="PLASTIC" & dominant_times_poison_executed == 0)$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="STATIC" & dominant_times_poison_executed == 0)$RANDOM_SEED)
  ),
  total=c(
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="NON-PLASTIC")$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="PLASTIC")$RANDOM_SEED),
    length(filter(summary_data, POISON_PENALTY==hitchiker_penalty & condition=="STATIC" )$RANDOM_SEED)
  ),
  condition=c(
    "NON-PLASTIC",
    "PLASTIC",
    "STATIC"
  )
)

ggplot(fig_data, aes(x=condition, y=num_execute_poison, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  geom_text(aes(label=num_execute_poison, y=num_execute_poison+0.25)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  )
```

Number of poison increases along lineage.

```{r}
poison_increases_fig <- ggplot(
    filter(summary_data, POISON_PENALTY==hitchiker_penalty),
    aes(x=condition, y=dominant_lineage_num_times_hitchhike_inst_exec_increases, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Increases in poison instruction execution along lineage",
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip()

poison_increases_fig
```

```{r}

poison_increases_per_gen_fig <- ggplot(
    filter(summary_data, POISON_PENALTY==hitchiker_penalty),
    aes(x=condition, y=dominant_lineage_num_times_hitchhike_inst_exec_increases_per_generation, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Increases in poison instruction execution along lineage (per generation)",
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip()

poison_increases_per_gen_fig
```

Pull it all together.

```{r}
# grid <- plot_grid(
#   dominant_extra_tasks_fig,
#   lineage_extra_tasks_discovered_fig + theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank()),
#   lineage_extra_tasks_lost_fig + theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank()),
#   nrow=1,
#   align="v",
#   labels="auto"
# )
# save_plot(
#    paste0(working_directory, "plots/", "complex-traits-panel.pdf"),
#    grid,
#    base_height=6,
#    base_asp=2.5
# )
# grid
```