# Evolution and maintenance of novel traits

The effect of adaptive phenotypic plasticity on the evolution and maintenance of novel tasks.

## Overview

```{r}
total_updates <- 200000
replicates <- 100

focal_traits <- c("not","nand","and","ornot","or","andnot")
traits_set_a <- c("not", "and", "or")
traits_set_b <- c("nand", "ornot", "andnot")
extra_traits <- c(
  "nor","xor","equals",
  "logic_3aa","logic_3ab","logic_3ac",
  "logic_3ad","logic_3ae","logic_3af",
  "logic_3ag","logic_3ah","logic_3ai",
  "logic_3aj","logic_3ak","logic_3al",
  "logic_3am","logic_3an","logic_3ao",
  "logic_3ap","logic_3aq","logic_3ar",
  "logic_3as","logic_3at","logic_3au",
  "logic_3av","logic_3aw","logic_3ax",
  "logic_3ay","logic_3az","logic_3ba",
  "logic_3bb","logic_3bc","logic_3bd",
  "logic_3be","logic_3bf","logic_3bg",
  "logic_3bh","logic_3bi","logic_3bj",
  "logic_3bk","logic_3bl","logic_3bm",
  "logic_3bn","logic_3bo","logic_3bp",
  "logic_3bq","logic_3br","logic_3bs",
  "logic_3bt","logic_3bu","logic_3bv",
  "logic_3bw","logic_3bx","logic_3by",
  "logic_3bz","logic_3ca","logic_3cb",
  "logic_3cc","logic_3cd","logic_3ce",
  "logic_3cf","logic_3cg","logic_3ch",
  "logic_3ci","logic_3cj","logic_3ck",
  "logic_3cl","logic_3cm","logic_3cn",
  "logic_3co","logic_3cp"
)

# Relative location of data.
working_directory <- "experiments/2021-01-31-complex-features/analysis/" # << For bookdown
# working_directory <- "./"
```

## Analysis dependencies

Load all required R libraries.

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(Hmisc)
library(boot)
source("https://gist.githubusercontent.com/benmarwick/2a1bb0133ff568cbe28d/raw/fb53bd97121f7f9ce947837ef1a4c65a73bffb3f/geom_flat_violin.R")
```

These analyses were conducted/knitted with the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
####### summary data #######
summary_data_loc <- paste0(working_directory, "data/aggregate.csv")
summary_data <- read.csv(summary_data_loc, na.strings="NONE")

summary_data$DISABLE_REACTION_SENSORS <- as.factor(summary_data$DISABLE_REACTION_SENSORS)
summary_data$chg_env <- summary_data$chg_env == "True"
summary_data$dominant_plastic_odd_even <- as.factor(summary_data$dominant_plastic_odd_even)
summary_data$sensors <- summary_data$DISABLE_REACTION_SENSORS == "0"
summary_data$is_plastic <- summary_data$dominant_plastic_odd_even == "True"
summary_data$extra_task_value <- as.factor(summary_data$extra_task_value)
summary_data <- filter(summary_data, extra_task_value == 0.1)

env_label_fun <- function(chg_env) {
  if (chg_env) {
    return("Fluctuating")
  } else {
    return("Constant")
  }
}

sensors_label_fun <- function(has_sensors) {
  if (has_sensors) {
    return("Sensors")
  } else {
    return("No sensors")
  }
}

condition_label_fun <- function(has_sensors, env_chg) {
  if (has_sensors && env_chg) {
    return("PLASTIC")
  } else if (env_chg) {
    return("NON-PLASTIC")
  } else {
    return("STATIC")
  }
}

summary_data$env_label <- mapply(
  env_label_fun,
  summary_data$chg_env
)
summary_data$sensors_label <- mapply(
  sensors_label_fun,
  summary_data$sensors
)
summary_data$condition <- mapply(
  condition_label_fun,
  summary_data$sensors,
  summary_data$chg_env
)

condition_order = c(
  "STATIC",
  "NON-PLASTIC",
  "PLASTIC"
)

###### time series #####
lineage_time_series_data_loc <- paste0(working_directory, "data/lineage_series.csv")
lineage_time_series_data <- read.csv(lineage_time_series_data_loc)

lineage_time_series_data$DISABLE_REACTION_SENSORS <- as.factor(lineage_time_series_data$DISABLE_REACTION_SENSORS)
lineage_time_series_data$chg_env <- lineage_time_series_data$chg_env == "True"
lineage_time_series_data$sensors <- lineage_time_series_data$DISABLE_REACTION_SENSORS == "0"
lineage_time_series_data$extra_task_value <- as.factor(lineage_time_series_data$extra_task_value)

lineage_time_series_data$env_label <- mapply(
  env_label_fun,
  lineage_time_series_data$chg_env
)
lineage_time_series_data$sensors_label <- mapply(
  sensors_label_fun,
  lineage_time_series_data$sensors
)
lineage_time_series_data$condition <- mapply(
  condition_label_fun,
  lineage_time_series_data$sensors,
  lineage_time_series_data$chg_env
)

####### misc #######
# Configure our default graphing theme
theme_set(theme_cowplot())
# Create directory to dump plots
dir.create(paste0(working_directory, "plots"), showWarnings=FALSE)
# Sample mean function
samplemean <- function(x, d) {
  return(mean(x[d]))
}
```

## The evolution of phenotypic plasticity

For sensor-enabled populations in fluctuating environments, we only transfered populations containing an optimally plastic genotype to phase two.

```{r, message=FALSE}
summary_data_grouped = dplyr::group_by(summary_data, sensors, env_label, condition, extra_task_value)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())

ggplot(summary_data_group_counts, aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  geom_text(aes(label=n, y=n+2)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  ylab("Number of replicates in phase two") +
  facet_wrap(~extra_task_value, labeller=label_both) +
  theme(
    legend.position="none"
  )
```

We can confirm our expectation that the dominant genotypes in non-plastic conditions are not phenotypically plastic.

```{r, message=FALSE}
summary_data_grouped = dplyr::group_by(summary_data, condition, is_plastic, extra_task_value)
summary_data_group_counts = dplyr::summarize(summary_data_grouped, n=dplyr::n())
ggplot(filter(summary_data_group_counts, is_plastic), aes(x=condition, y=n, fill=condition)) +
  geom_col(position=position_dodge(0.9)) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  ylim(0, 100) +
  geom_text(aes(label=n, y=n+1)) +
  ylab("Number of replicates with a plastic dominant genotype") +
  facet_wrap(~extra_task_value, labeller=label_both) +
  theme(
    legend.position="none"
  )
```

## Novel task performance - final dominant genotype

How many novel tasks do final dominant genotypes perform?

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_extra_tasks, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Novel tasks performed by final dominant") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-extra-tasks.pdf"),
    width=15,
    height=10
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_extra_tasks)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$dominant_extra_tasks)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_extra_tasks)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=dominant_extra_tasks~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$dominant_extra_tasks,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

## Novel task performance - final population

How many novel tasks are performed across the final population (1% of organisms must perform to count)?

```{r}
ggplot(summary_data, aes(x=condition, y=final_pop_extra_tasks_0.01, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$final_pop_extra_tasks_0.01)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$final_pop_extra_tasks_0.01)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$final_pop_extra_tasks_0.01)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=final_pop_extra_tasks_0.01~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$final_pop_extra_tasks_0.01,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

## Novel task discovery - population

```{r}
ggplot(summary_data, aes(x=condition, y=discovered_extra_tasks_0.01, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$discovered_extra_tasks_0.01)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$discovered_extra_tasks_0.01)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$discovered_extra_tasks_0.01)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=discovered_extra_tasks_0.01~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$discovered_extra_tasks_0.01,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

## Novel task discovery - lineage of final dominant genotype

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_extra_traits_discovered, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Extra tasks discovered along dominant lineage") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-lineage-extra-tasks-discovered.pdf"),
    width=15,
    height=10
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_discovered)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_extra_traits_discovered)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_discovered)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=dominant_lineage_extra_traits_discovered~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$dominant_lineage_extra_traits_discovered,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

### Per-generation novel task discovery - lineage of final dominant genotype

```{r}
summary_data$dominant_lineage_extra_traits_discovered_per_generation <- summary_data$dominant_lineage_extra_traits_discovered / summary_data$dominant_generation_born
ggplot(summary_data, aes(x=condition, y=dominant_lineage_extra_traits_discovered_per_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Lineage task discovery (per generation)") +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  ggsave(paste0(working_directory, "plots/", "discovery-per-generation.png"))
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_discovered_per_generation)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_extra_traits_discovered_per_generation)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_discovered_per_generation)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=dominant_lineage_extra_traits_discovered_per_generation~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$dominant_lineage_extra_traits_discovered_per_generation,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

## Novel tasks gained along final dominant lineages

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_extra_traits_gained, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Extra tasks gained along dominant lineage") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-lineage-extra-tasks-gained.pdf"),
    width=15,
    height=10
  )
```

## Novel task loss along final dominant lineages

```{r}
ggplot(summary_data, aes(x=condition, y=dominant_lineage_extra_traits_lost, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Extra tasks lost along dominant lineage") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  ) +
  ggsave(
    paste0(working_directory, "plots/dominant-lineage-extra-tasks-lost.pdf"),
    width=15,
    height=10
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=dominant_lineage_extra_traits_lost~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$dominant_lineage_extra_traits_lost,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

### Per-generation novel task loss along final dominant lineages

```{r}
summary_data$dominant_lineage_extra_traits_lost_per_generation <- summary_data$dominant_lineage_extra_traits_lost / summary_data$dominant_generation_born
ggplot(summary_data, aes(x=condition, y=dominant_lineage_extra_traits_lost_per_generation, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Extra tasks lost along dominant lineage (per generation)") +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  ) +
  ggsave(paste0(working_directory, "plots/", "tasks-lost-per-generation.png"))
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost_per_generation)
)
paste0(
  "STATIC median: ",
  median(filter(summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost_per_generation)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost_per_generation)
)

reward_level <- 0.1
dom_task_data <- filter(summary_data, extra_task_value==reward_level)
kruskal.test(
  formula=dominant_lineage_extra_traits_lost_per_generation~condition,
  data=dom_task_data
)
pairwise.wilcox.test(
  x=dom_task_data$dominant_lineage_extra_traits_lost_per_generation,
  g=dom_task_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

## How many instances of novel trait loss co-occurred with changes in base phenotype?

Task loss linked with primary trait changes.

```{r}
lost_traits_summary_data <- filter(summary_data, extra_task_value==0.1 & dominant_lineage_extra_traits_lost>0)
lost_traits_summary_data$frac_linked_extra_trait_loss <- lost_traits_summary_data$dominant_lineage_extra_traits_lost_linked_to_primary_change / lost_traits_summary_data$dominant_lineage_extra_traits_lost

# ggplot(lost_traits_summary_data, aes(x=frac_linked_extra_trait_loss, fill=condition)) +
#   geom_density() +
#   facet_grid(
#     condition~extra_task_value,
#     labeller=label_both
#   ) +
#   theme(
#     legend.position="none"
#   ) +
#   scale_fill_brewer(
#     palette="Paired"
#   ) +
#   scale_color_brewer(
#     palette="Paired"
#   ) +
#   ggsave(
#     paste0(working_directory, "plots/dominant-lineage-extra-tasks-lost-linkage.png"),
#     width=15,
#     height=10
#   )
```

```{r}
ggplot(lost_traits_summary_data, aes(x=condition, y=frac_linked_extra_trait_loss, fill=condition)) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  facet_wrap(
    ~extra_task_value,
    labeller=label_both
  ) +
  theme(
    legend.position="none"
  )
```

```{r}
paste0(
  "PLASTIC median: ",
  median(filter(lost_traits_summary_data, condition=="PLASTIC")$frac_linked_extra_trait_loss)
)
paste0(
  "STATIC median: ",
  median(filter(lost_traits_summary_data, condition=="STATIC")$frac_linked_extra_trait_loss)
)
paste0(
  "NON-PLASTIC median: ",
  median(filter(lost_traits_summary_data, condition=="NON-PLASTIC")$frac_linked_extra_trait_loss)
)

kruskal.test(
  formula=frac_linked_extra_trait_loss~condition,
  data=lost_traits_summary_data
)
pairwise.wilcox.test(
  x=lost_traits_summary_data$frac_linked_extra_trait_loss,
  g=lost_traits_summary_data$condition,
  p.adjust.method="bonferroni",
  conf.int=TRUE,
  conf.level=0.95
)
```

```{r}
sum(filter(lost_traits_summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change)
sum(filter(lost_traits_summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_nonplastic <- sum(filter(lost_traits_summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change) / sum(filter(lost_traits_summary_data, condition=="NON-PLASTIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_nonplastic

sum(filter(lost_traits_summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change)
sum(filter(lost_traits_summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_plastic <- sum(filter(lost_traits_summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change) / sum(filter(lost_traits_summary_data, condition=="PLASTIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_plastic

sum(filter(lost_traits_summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change)
sum(filter(lost_traits_summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_nonplastic <- sum(filter(lost_traits_summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost_linked_to_primary_change) / sum(filter(lost_traits_summary_data, condition=="STATIC")$dominant_lineage_extra_traits_lost)
aggregate_frac_linked_extra_trait_loss_nonplastic
```

## Manuscript figures

### Overall task performance

```{r}
extra_task_reward_value=0.1
dominant_extra_tasks_fig <- ggplot(
    filter(summary_data, extra_task_value==extra_task_reward_value),
    aes(x=condition, y=dominant_extra_tasks, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Final dominant novel traits"
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip()
dominant_extra_tasks_fig
```

### Overall task discovery

```{r}
lineage_extra_tasks_discovered_fig <- ggplot(
    filter(summary_data, extra_task_value==extra_task_reward_value),
    aes(x=condition, y=dominant_lineage_extra_traits_discovered, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Novel traits discovered on lineage"
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip()
lineage_extra_tasks_discovered_fig
```

### Overall task loss

```{r}
lineage_extra_tasks_lost_fig <- ggplot(
    filter(summary_data, extra_task_value==extra_task_reward_value),
    aes(x=condition, y=dominant_lineage_extra_traits_lost, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order,
    labels=condition_order
  ) +
  scale_y_continuous(
    name="Novel traits lost on lineage (log scale)",
    trans="pseudo_log",
    breaks=c(0,10,100,1000),
    limits=c(-1,1000)
  ) +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  theme(
    legend.position="none"
  ) +
  coord_flip()

lineage_extra_tasks_lost_fig
```


### Per-generation task discovery

```{r}
lineage_per_gen_task_discovery <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_extra_traits_discovered_per_generation, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Per-generation task discovery") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  )
lineage_per_gen_task_discovery
```

### Per-generation task loss

```{r}
lineage_per_gen_task_loss <- ggplot(
    summary_data,
    aes(x=condition, y=dominant_lineage_extra_traits_lost_per_generation, fill=condition)
  ) +
  geom_flat_violin(
    position = position_nudge(x = .2, y = 0),
    alpha = .8
  ) +
  geom_point(
    mapping=aes(color=condition),
    position = position_jitter(width = .15),
    size = .5,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = .1,
    outlier.shape = NA,
    alpha = 0.5
  ) +
  scale_x_discrete(
    name="Condition",
    limits=condition_order
  ) +
  ylab("Per-generation task loss") +
  scale_fill_brewer(
    palette="Paired"
  ) +
  scale_color_brewer(
    palette="Paired"
  ) +
  coord_flip() +
  theme(
    legend.position="none"
  )
lineage_per_gen_task_loss
```

## Combined panel

```{r}
grid <- plot_grid(
  dominant_extra_tasks_fig,
  lineage_extra_tasks_discovered_fig + theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank()),
  lineage_extra_tasks_lost_fig + theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank()),
  NULL,
  lineage_per_gen_task_discovery,
  lineage_per_gen_task_loss + theme(axis.ticks.y=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank()),
  nrow=2,
  align="v",
  # labels="auto"
  labels=c("a", "b","c","", "d","e")
)
save_plot(
   paste0(working_directory, "plots/", "complex-traits-panel.pdf"),
   grid,
   base_height=12,
   base_asp=3/2
)
grid
# save_plot(
#    paste0(working_directory, "plots/", "complex-traits-panel.png"),
#    grid,
#    base_height=6,
#    base_asp=2.5
# )
```
