---
title: "2020-09-03 Experiment Analysis"
output:
  html_document:
    keep_md: no
    toc: true
    toc_float: true
    toc_depth: 4
    collapsed: false
    theme: default
    code_folding: hide
  pdf_document:
    toc: true
    toc_depth: 4
---

## Analysis Dependencies

```{r, message=FALSE}
library(ggplot2)   # (Wickham, 2016)
library(tidyr)     # (Wickham and Henry, 2020)
library(dplyr)     # (Wickham et al., 2020)
library(reshape2)  # (Wickham, 2007)
library(cowplot)   # (Wilke, 2019)
library(patchwork) # (Pederson, 2020)
library(viridis)   # (Garnier, 2018)
library(hexbin)
```

We conducted these analyses using the following computing environment:

```{r}
print(version)
```

## Setup

```{r}
data_path <- "./data/aggregate.csv"
agg_data <- read.csv(data_path, na.strings="NONE")

agg_data$DISABLE_REACTION_SENSORS <- as.factor(agg_data$DISABLE_REACTION_SENSORS)
agg_data$environment = sub("_.*", "", agg_data$environment)
agg_data$environment <- factor(
  agg_data$environment,
  levels=c("a-u0", "b-u0", "all-u0",
            "chg-u3", "chg-u10","chg-u30",
            "chg-u100","chg-u300","chg-u1000"))
agg_data$chg_env <- as.factor(agg_data$chg_env)
agg_data$plastic_odd_even <- as.factor(agg_data$plastic_odd_even)
agg_data$sensors <- agg_data$DISABLE_REACTION_SENSORS == "0"

theme_set(theme_cowplot())
color_true = '#8da0cb'
color_false = '#fc8d62'
```

## Graphs!

### Number of phase 2 replicates
```{r}
data_grouped = dplyr::group_by(agg_data, sensors, environment)
data_summary = dplyr::summarize(data_grouped, n = dplyr::n())
ggplot(data_summary, aes(x=sensors, y = n, fill = sensors)) +
  geom_col(position=position_dodge(0.9)) + 
  geom_text(aes(label = n, y = n + 4)) + 
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_fill_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  ylab("Replicates") +
  ylim(0, 105) +
  facet_wrap(~environment, nrow=1) +
  ggtitle("Number of replicates in phase two") +
  ggsave("num_replicates.pdf", width=10, height=7)
```

### Plasticity

```{r}
ggplot(filter(agg_data, plastic_odd_even=="True"), aes(x=sensors, fill = sensors)) +
  geom_bar(position=position_dodge(0.9)) +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9), vjust=0
  )  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_fill_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  ylab("plastic final dominant (odd/even environment)") +
  ylim(0, 50) +
  facet_wrap(~environment, nrow=1) +
  ggtitle("Plastic?") +
  ggsave("dom_plastic.pdf", width=10, height=7)
```

```{r}
ggplot(filter(agg_data, optimal_plastic=="True"), aes(x=sensors, fill = sensors)) +
  geom_bar(position=position_dodge(0.9)) +
  geom_text(
    stat="count",
    mapping=aes(label=..count..),
    position=position_dodge(0.9), vjust=0
  )  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_fill_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  ylab("plastic final dominant (odd/even environment)") +
  ylim(0, 50) +
  facet_wrap(~environment, nrow=1) +
  ggtitle("Optimal plastic?") +
  ggsave("dom_plastic_optimal.pdf")
```

### Match scores

```{r}
ggplot(filter(agg_data, chg_env=="True"), aes(x=sensors, y=match_score_odd_even, color = sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, nrow=1) 
```

### MRCA Changes

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_mrca_changes, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, nrow=1) +
  ggsave("mrca_changes.pdf", width=10, height=7)
```

### Lineage Length (genotypes)

```{r}
ggplot(agg_data, aes(x=sensors, y=lineage_length_genotypes, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, nrow=1) +
  # theme(
  #   panel.border = element_rect(colour = "black", fill=NA, size=1)
  # ) +
  ggsave("lineage_length_genotypes.pdf", width=10, height=7)
```

### Mutation Accumulation

```{r}
ggplot(agg_data, aes(x=sensors, y=total_mut_cnt, color=sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, ncol=1, scales="free") +
  ggsave("total_mutations_accumulated.pdf", width=5, height=20)
```

### Phylogenetic Diversity

Current phylogentic diversity = size of phylogentic tree (number of active taxa + ancestral taxa represented in phylogeny).

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_current_phylogenetic_diversity, color = sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, nrow=1) 
```

### Taxa Entropy (diversity)

Extant diversity measured as entropy of taxa in population.

```{r}
ggplot(agg_data, aes(x=sensors, y=phylo_diversity, color = sensors)) +
  geom_boxplot()  +
  scale_x_discrete(name="Sensors", limits=c(TRUE, FALSE), labels=c("T","F")) +
  scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
  facet_wrap(~environment, nrow=1) 
```


## Diversity over time data

```{r}

phylo_diversity_path <- "./data/phylodiversity_ot.csv"
phylo_ot_data <- read.csv(phylo_diversity_path, na.strings="NONE")

phylo_ot_data$DISABLE_REACTION_SENSORS <- as.factor(phylo_ot_data$DISABLE_REACTION_SENSORS)
phylo_ot_data$environment = sub("_.*", "", phylo_ot_data$environment)
phylo_ot_data = phylo_ot_data[phylo_ot_data$environment != 0,] # Straggler data from trial run
phylo_ot_data$environment <- factor(
  phylo_ot_data$environment,
  levels=c("a-u0", "b-u0", "all-u0",
           "chg-u3", "chg-u10","chg-u30",
           "chg-u100","chg-u300","chg-u1000")
)
phylo_ot_data$chg_env <- as.factor(phylo_ot_data$chg_env)
# phylo_ot_data$plastic_odd_even <- as.factor(phylo_ot_data$plastic_odd_even)
phylo_ot_data$sensors <- phylo_ot_data$DISABLE_REACTION_SENSORS == "0"

phylo_ot_data$diversity <- as.numeric(as.character(phylo_ot_data$diversity))
phylo_ot_data$variance_pairwise_distance <- as.numeric(as.character(phylo_ot_data$variance_pairwise_distance))

```

```{r}
graph_diversity_ot <- function(env) {
  ot_data <- filter(phylo_ot_data, environment==env)
  
  ggplot(
    ot_data, 
    aes(x=update, y=diversity, fill=sensors, color=sensors)
    ) +
    stat_summary(fun="mean", geom="line") +
    stat_summary(fun.data="mean_cl_boot", 
                 fun.args=list(conf.int=0.95), 
                 geom="ribbon", 
                 alpha=0.2,
                 linetype=0) +
    ggtitle(env) +
    scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
    ggsave(paste("diversity_ot_",env,".pdf", sep=""), width=20, height=10)
  
  ggplot(
    ot_data, 
    aes(x=update, y=mean_pairwise_distance, fill=sensors, color=sensors)
    ) +
    stat_summary(fun="mean", geom="line") +
    stat_summary(fun.data="mean_cl_boot", 
                 fun.args=list(conf.int=0.95), 
                 geom="ribbon", 
                 alpha=0.2,
                 linetype=0) +
    ggtitle(env) +
    scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
    ggsave(paste("mean_pairwise_dist_ot_",env,".pdf", sep=""), width=20, height=10)
  
  ggplot(
    ot_data, 
    aes(x=update, y=sum_pairwise_distance, fill=sensors, color=sensors)
    ) +
    stat_summary(fun="mean", geom="line") +
    stat_summary(fun.data="mean_cl_boot", 
                 fun.args=list(conf.int=0.95), 
                 geom="ribbon", 
                 alpha=0.2,
                 linetype=0) +
    ggtitle(env) +
    scale_color_manual(values = c("FALSE" = color_false, "TRUE" = color_true)) + 
    ggsave(paste("sum_pairwise_dist_ot_",env,".pdf", sep=""), width=20, height=10)

}

graph_diversity_ot("chg-u300")
graph_diversity_ot("chg-u1000")
graph_diversity_ot("chg-u10")
graph_diversity_ot("chg-u100")
# "a-u0", "b-u0", "all-u0",
#            "chg-u3", "chg-u10","chg-u30",
#            "chg-u100","chg-u300","chg-u1000"

#ggplot(
#  filter(ot_data, environment=="chg-u100" & update <= 199000), 
#  aes(x=update, y=variance_pairwise_distance, fill=sensors, color=sensors)
#  ) +
#  stat_summary(fun="mean", geom="line") +
#  stat_summary(fun.data="mean_cl_boot", 
#               fun.args=list(conf.int=0.95), 
#               geom="ribbon", 
#               alpha=0.2,
#               linetype=0) +
#  ggtitle("chg-u100") +
#  ggsave("variance_pairwise_dist_ot_chg-u100.pdf", width=20, height=10)
```